Survey of Cache Oblivious B-trees

### 2011-12-12 paper.tex

In order for B-trees to remain versatile, they must be able to accommodate
various forms of data. However, traditional B-trees and string B-trees have
suboptimal performance when variable length strings are used as keys. In
addition, string B-trees do not support compression, and commonly used front
compression is handled sub-optimally. Regular B-trees and string B-trees are
not cache oblivious, and thus incur poorer performance as described previously.
In \cite{BenderFaKu06}, Bender et al. address these issues by designing a Cache
Oblivious String B-tree (COSB-tree), which they argue works efficiently for
variable length string keys. They also provide a modified front-compression
scheme, which gives their approach the advantages of using less space and
having improved locality within the compressed data.

>>>>

In order for B-trees to remain versatile, they must be able to accommodate
various forms of data. Bender et al. assert the following weaknesses in existing string B-tree designs: traditional B-trees have
suboptimal performance when variable length strings are used as keys\cite{BenderDemainColton}. String B-trees\cite{Ferragina98} handle variable-length strings well but do not use compression. Front compression\cite{BayerUn77,ClarkSaSt69,Comer79,Wagner73} is used in many B-trees, but is handled sub-optimally\cite{BenderFaKu06}. Regular B-trees and string B-trees are
not cache oblivious, and thus incur poorer performance as described
previously by Bender et al.\cite{BenderDemainColton}. Range queries have poor performance in B-trees\cite{BenderFaKu06}.

In \cite{BenderFaKu06}, Bender et al. address these issues by designing a Cache
Oblivious String B-tree (COSB-tree), which they argue works efficiently in
these problem areas for variable length string keys. They also provide a
modified front-compression scheme, which gives their approach the advantages
of using less space and having improved locality within the compressed data.

\subsection{Description}
Bender et al. perform 2 experiments in \cite{BenderFaKu06}, in which a regular B-tree implementation tuned to various block sizes is compared to a static COSB-tree implementation. The experiment involved measuring the amount of time required to perform: 440,000 and 450,000 inserts of random values, a range query of all values, and 1000 random searches. The results of the experiment show that COSB-trees
outperformed B-trees most of the time for random searches, range queries and
random inserts. However, some point between 440,000 and 450,000 inserts the
COSB-tree had to rebalance the entire tree, causing it to fall behind in this
measurement. However, Bender et al. argue that the experiment was biased
towards the B-tree implementation, since the trees are "young", and thus still
well organized. The COSB-tree implementation is also compared to Berkeley DB
\cite{BerkeleyDB}, and achieves performance comparable to Berkeley DB with
tuned parameters. The results seem to show that there is a window in which
large and small-block size B-trees perform poorly, and that COSB-trees may be
a reasonable alternative to improve performance.
